\begin{frontmatter}              % The preamble begins here.


%\pretitle{Pretitle}
\title{Chemical Identity Lost in Regulation\@:\\A Study of Semantic Interoperability in European Chemical Substance Data}
\runningtitle{Chemical Identity Lost in Regulation}
%\subtitle{Subtitle}

%% \\ifblindreview\\else[\s\S]*?\\fi
\ifblindreview\else
\author[A]{\fnms{Geert} \snm{Van Haute}\orcid{0009-0002-9836-8139}%
\thanks{Corresponding Author: Geert Van Haute, contact details.}},
\author[A]{\fnms{Pieter} \snm{Fannes}\orcid{....-....-....-....}}
,
\author[A]{\fnms{Stijn} \snm{Goedertier}\orcid{....-....-....-....}}
and
\author[A]{\fnms{Maxim} \snm{Van de Wynckel}\orcid{0000-0003-0314-7107}}

\runningauthor{G. Van Haute et al.}
\address[A]{Department of Environment and Spatial Development, Flemish Government, Belgium}
\fi


\begin{abstract}
Chemical substance identity is inconsistently represented across European regulatory datasets, creating fundamental barriers to semantic interoperability. While structure-based identifiers such as InChI and InChIKey unambiguously define chemical entities, regulatory systems rely on heterogeneous identifiers, including CAS numbers, EC numbers, textual names, and group descriptions. These identifiers often refer to mixtures, substance groups, or analytical parameters rather than uniquely defined molecular structures.

In this study, we analyse interoperability challenges across 18 European regulatory datasets from ECHA and related sources. Where possible, we map entries to InChIKeys, classify records by structural linkability, assess the consistency of CAS number mappings, evaluate embedding-based semantic matching, estimate the effort required for manual reconciliation, and implement a knowledge-graph-based integration approach.

We find that only 62.5\% of entries can be directly linked to a defined chemical structure, leaving 37.5\% inherently non-structure-based. CAS numbers exhibit both one-to-many and many-to-one mappings with molecular structures, undermining their role as unique identifiers. Embedding-based approaches fail to recover chemical identity for non-structure substances, and the best model achieves a ChemOnt Hit@1 of only 5.75\%. The combinatorial complexity of cross-dataset identifier mappings makes retrospective harmonisation infeasible at scale.

These findings demonstrate that semantic interoperability in chemical regulation cannot be achieved through post hoc technical solutions alone. Instead, it must be addressed at the level of data modelling and legislation, by adopting structure-based identifiers, formal ontologies, and machine-readable representations embedded directly in regulatory data standards. To support further research and reuse, we provide scripts for constructing an integrated knowledge graph spanning the analysed regulatory sources.


\end{abstract}



\begin{keyword}
semantic interoperability\sep regulatory interoperability\sep Interoperable Europe\sep chemical substance identification 
\end{keyword}
\end{frontmatter}
