\section{Methods}
\label{sec:methods}

\subsection{Data Sources and Integration}
\label{sec:datasources}

Substance records were compiled from multiple regulatory lists published through
the European Chemicals Agency (ECHA), including the candidate list of substances
of very high concern (SVHC), the authorisation list, the restriction list, the
persistent organic pollutants (POPs) list, the harmonised classification and
labelling list, the EU positive list, and the SVHC identification list, as well
as the EU Pesticides Database and the Flemish administrative
\textit{sommatie\_stoffen} list.  For each substance record, the available
identifiers were collected: CAS registry number, EC number, and substance name.

Where a CAS number was present, the corresponding InChIKey was retrieved via the
PubChem PUG-REST API~\cite{Kim2015}; where a CAS lookup yielded no result, a
fallback lookup by substance name was attempted.  All API responses were cached
locally using the MD5 hash of the query as the filename, to avoid redundant
network requests.  A delay of 0.2~s was enforced between successive requests to
comply with the PubChem rate limit.  Known InChIKeys were subsequently propagated
across all records sharing the same substance name within the dataset.  The
resulting dataset was stored as a unified R~\cite{RCoreTeam2024} data frame
linking each regulatory entry to its source list and all available identifiers
(\texttt{data/processed/all\_substances.rds}).  RDF serialisation and graph
merging were performed using Apache Jena~5.6.0.

\subsection{Chemical Identity Classification}
\label{sec:classification}

To characterise the heterogeneity of regulatory substance entries, a rule-based
classification scheme was applied to assign each unique substance name to one of
seven mutually exclusive entity types.  Classification was implemented as an
ordered conditional lookup in which the \textit{Molecule} type was evaluated
first, so that the availability of a structure-based identifier always takes
precedence over textual pattern matching.  The seven types are:

\begin{enumerate}
  \item \textit{Molecule} --- any entry for which a valid InChIKey is available,
    indicating that the substance can be represented as a defined molecular
    structure;
  \item \textit{Substance group} --- entries whose names contain phrases such as
    ``and its compounds'', ``and its salts'', ``and its esters or isomers'', or
    group acronyms (e.g.\ PFAS, PCB, PAH, PBDE, HBCDD);
  \item \textit{Analytical parameter} --- entries describing measurement
    quantities rather than individual substances, identifiable by phrases such
    as ``Reaction mass of\ldots'', carbon-range notation ($\mathrm{C}_n$--$\mathrm{C}_m$),
    or semicolon-separated compound lists;
  \item \textit{Mixture} --- entries corresponding to complex compositions of
    unknown or variable structure, including trade-name products, petroleum
    derivatives (naphtha, kerosine, petrolatum), and creosote;
  \item \textit{Regulatory entry} --- administrative entries referencing
    legislative artefacts rather than chemical entities, such as entries
    formatted as ``Entry~$N$'', references to EU directives, or annex
    descriptions;
  \item \textit{CAS without structure} --- entries with a CAS number for which
    no InChIKey could be retrieved, indicating either that the substance is not
    indexed in PubChem or that it represents an incompletely characterised
    (UVCB) substance;
  \item \textit{Unclassified} --- entries with neither an InChIKey nor a CAS
    number and not matching any of the above patterns.
\end{enumerate}

\subsection{Linkability Taxonomy}
\label{sec:linkability}

Building on the entity-type classification, a seven-tier linkability taxonomy
was derived to express the degree to which a regulatory substance entry can be
unambiguously associated with a defined chemical structure.  For substance-group
entries, the presumed base-compound name was extracted by regular expression and
an InChIKey lookup via PubChem was attempted, with a 0.2~s delay between
requests, yielding a distinction between groups whose parent compound is
structure-resolvable and those for which it is not.  Analytical parameters and
mixtures were consolidated into a single UVCB/Mixture tier.  The taxonomy tiers,
in decreasing order of chemical specificity, are:
(1)~Structure (InChIKey);
(2)~Group --- base compound resolvable;
(3)~Group --- base compound not resolvable;
(4)~CAS without structure;
(5)~UVCB\,/\,Mixture;
(6)~Regulatory entry;
(7)~Unidentified.

This taxonomy was serialised as SKOS Collections in RDF Turtle format, with
substance URIs minted as
\texttt{https://data.omgeving.vlaanderen.be/id/concept/chemical\_sub\-stance/\{md5\}}.
The classified dataset was saved as \texttt{data/processed/all\_substances\_classified.rds}
for use in downstream analyses.

\subsection{Identifier Mapping Consistency}
\label{sec:identifier-mapping}

For all records possessing both a CAS number and an InChIKey, bidirectional
mapping consistency was assessed in two directions.  First, the number of
distinct InChIKeys associated with each CAS number was computed: a CAS number
mapping to more than one InChIKey indicates that the same registry identifier has
been assigned to multiple molecular structures, for example due to differing
stereochemical representations, hydrates, or salt forms.  Second, the number of
distinct CAS numbers associated with each InChIKey was computed: an InChIKey
associated with multiple CAS numbers reflects the common regulatory practice of
assigning separate registry identifiers to the same molecular structure in
different trade or regulatory contexts.  Both distributions were visualised on a
$\log_{10}$ frequency axis~\cite{Cleveland1993} to reveal the long tail of the
distribution.  These analyses motivate the use of InChIKey, rather than CAS
number, as the primary linking key throughout all subsequent cross-list analyses.

\subsection{Structure Coverage and Linking Ratio}
\label{sec:coverage}

For each regulatory source, the \textit{linking ratio} was defined as the number
of unique InChIKeys divided by the total number of regulatory records in that
source.  A ratio of~1 indicates that every record corresponds to a distinct
molecular structure; a ratio below~1 reflects structural redundancy (multiple
records per structure), a high non-structure fraction, or both.  The linking
ratio was visualised as a $\log_{10}$--$\log_{10}$ scatter plot with a~1:1
reference diagonal~\cite{Cleveland1993,Tufte2001}, enabling comparison across
sources spanning three orders of magnitude in size.  This analysis provides the
per-source structural density estimate that feeds the workload model described in
Section~\ref{sec:workload}.

\subsection{Cross-List Substance Overlap}
\label{sec:overlap}

Four complementary analyses characterised how structure-defined substances are
distributed across regulatory sources.

\textit{UpSet plots}~\cite{Lex2014} were used to visualise multi-set
intersections of structure-defined substances across all sources with at least
ten unique InChIKeys; the threshold of ten was chosen as a display heuristic to
suppress near-empty sets that produce unreadable intersection bars.  The 30
largest intersections were shown, as the full $2^n - 1$ possible intersections
are not interpretable for $n > 5$.

Pairwise Jaccard similarity was computed for all source pairs:
\begin{equation}
  J(A,B) = \frac{|A \cap B|}{|A \cup B|},
  \label{eq:jaccard}
\end{equation}
where $A$ and $B$ denote the sets of unique InChIKeys in two sources,
respectively.  Only structure-defined entries were included.  The resulting
similarity matrix was subjected to hierarchical clustering using average linkage
on the complementary dissimilarity $1 - J$~\cite{Ward1963,Murtagh2014} and
visualised as a clustered heatmap; text labels were suppressed for $J < 0.02$
to avoid visual noise from near-zero overlaps.

A second UpSet plot was constructed restricted to the seven principal regulatory
obligation lists: the restriction list, candidate list, authorisation list, POPs
list, EU positive list, harmonised classification list, and SVHC identification
list.

Finally, a force-directed network~\cite{Fruchterman1991} was constructed in
which nodes represent regulatory sources (node size $\propto \log_{10}(\text{list
size} + 1)$) and edges connect pairs of sources sharing at least five unique
InChIKeys (edge width $\propto \log_{10}(\text{shared InChIKeys} + 1)$).  The
threshold of five shared InChIKeys suppresses edges representing negligible
structural overlap that would otherwise clutter the layout.

\subsection{Semantic Embedding and Clustering of Non-Structure Substances}
\label{sec:clustering}

Substance entries that could not be resolved to a chemical structure --- excluding
administrative regulatory entries --- were analysed through unsupervised text
embedding and clustering.  Each substance name was encoded using the
\texttt{all-MiniLM-L6-v2} sentence-transformer model~\cite{Reimers2019,Wang2020MiniLM},
yielding a 384-dimensional dense vector representation.

The optimal number of clusters was determined by computing the silhouette
score~\cite{Rousseeuw1987} for k-means solutions with $k \in \{2, \ldots, 12\}$
(25 random starts, maximum 100 iterations, random seed~42).  The optimal
solution ($k = 6$) was applied using Lloyd's algorithm~\cite{Lloyd1982} to
partition the substance names into six semantically coherent groups.

For visualisation, the embeddings were projected to two dimensions using
UMAP~\cite{McInnes2018} (15 nearest neighbours, minimum distance~0.1, cosine
metric, seed~42).

The six resulting clusters were labelled manually upon visual inspection of the
projections:
(1)~reaction masses;
(2)~complex substances (UVCB with partial structure);
(3)~petroleum and coal tar fractions;
(4)~inorganic compounds and metal salts;
(5)~trade names, codes and biological materials;
(6)~polymers, fatty acids and surfactants.

\subsection{Chemical Taxonomy Matching}
\label{sec:chemont}

Structure-defined and non-structure substances were mapped to standardised
chemical classes using two complementary approaches.

For \textit{non-structure substances}, the cosine similarity~\cite{Manning2008}
was computed between the sentence-transformer embeddings of each substance name
and the embeddings of all class labels in the ChemOnt~2.1 ontology~\cite{DjoumbouFeunang2016}.
Substance names matching patterns characteristic of reaction masses, petroleum
fractions, or UVCB descriptions were excluded prior to matching, as these names
do not correspond to a single chemical class.  For each remaining substance, the
top-3 most similar ChemOnt classes were identified and a composite confidence
score was computed as
\begin{equation}
  c = 0.6 \cdot s_1 + 0.3 \cdot \Delta_{12} + 0.1 \cdot \Delta_{23},
  \label{eq:confidence}
\end{equation}
where $s_1$ is the cosine similarity of the best-matching class and $\Delta_{12}$,
$\Delta_{23}$ are the score gaps between the first and second, and second and
third candidates, respectively.  Retaining three candidates rather than one is
necessary to compute both gap metrics.  High-quality matches were retained subject
to the empirically determined thresholds $s_1 > 0.8$ and $\Delta_{12} > 0.05$,
selected by sweeping the primary threshold between 0.3 and~0.9 and inspecting
the resulting quality landscape.

For \textit{structure-defined substances}, the direct-parent ChemOnt class was
retrieved for each unique InChIKey via the ClassyFire REST API~\cite{DjoumbouFeunang2016},
with all API responses cached locally to avoid redundant requests.

\subsection{Knowledge Graph Construction}
\label{sec:kg}

All processed substance data --- including the linkability taxonomy
(Section~\ref{sec:linkability}), cluster assignments
(Section~\ref{sec:clustering}), and ClassyFire annotations
(Section~\ref{sec:chemont}) --- were serialised to RDF Turtle format.  Substance
URIs were minted using MD5 hashes of InChIKeys for structure-defined entries, or
of substance names otherwise, to ensure stable and opaque identifiers.  The
linkability taxonomy and cluster assignments were represented as
\texttt{skos:Collection} instances~\cite{W3C2009SKOS}, with \texttt{skos:narrower}
triples encoding group-to-base-compound parent relationships.  Provenance
annotations were recorded using the OA Annotation Ontology~\cite{W3C2017OA}.

All component graphs were merged with the ChemOnt~2.1 SKOS ontology using
Apache Jena~5.6.0 \texttt{riot}, with semantic enrichment performed via SPARQL
CONSTRUCT queries.

To integrate biological hazard evidence, the ChEBI ontology~\cite{DjoumbouFeunang2016}
was linked to the knowledge graph through a four-step pipeline implemented as a
companion shell script.  First, the ChEBI OWL ontology was downloaded from the
European Bioinformatics Institute and converted to RDF Turtle using Jena
\texttt{riot}.  Second, the converted ontology was merged with the project's
substance taxonomy graph using Jena.  Third, a SPARQL CONSTRUCT query
(\texttt{exact\_match\_chebi.rq}) linked project substance concepts to their
corresponding ChEBI entities via InChIKey string matching
(\texttt{skos:exactMatch}), using the ChEBI \texttt{chemrof:inchi\_key\_string}
property as the join key.  Fourth, a second SPARQL CONSTRUCT query
(\texttt{chebi\_annotaties.rq}) traversed the ChEBI role hierarchy via the
\texttt{RO:0000087} (\textit{has role}) object property and materialised each
biological role assignment as an OA Annotation triple~\cite{W3C2017OA} linking
the project substance concept to the corresponding ChEBI role class.  Finally,
a SPARQL SELECT query with \texttt{GROUP\_CONCAT} (\texttt{chebi\_to\_table.rq})
exported a flat CSV table (\texttt{data/processed/chebi.csv}) containing, per
substance, pipe-separated lists of biological activity labels and ECHA list
memberships.  This table is the primary input for the composite priority scoring
described in Section~\ref{sec:prioritisation}.  The unified knowledge graph was
stored at \texttt{data/processed/rdf/substances\_taxonomy.ttl}.

\subsection{Workload Estimation for Manual Curation}
\label{sec:workload}

Regulatory entries that cannot be resolved to a chemical structure must be
related to one another through manual expert assessment.  For each pair of
non-structure entries, an expert must determine which of four mutually exclusive
semantic relations holds: (1)~\textit{equivalence} --- both entries refer to the
same set of substances; (2)~\textit{subset} --- one entry is fully contained
within the other; (3)~\textit{overlap} --- the entries share some but not all
substances; or (4)~\textit{disjoint} --- the entries share no substances.  Only
by establishing these relations explicitly can the semantic alignment required
for cross-list interoperability be achieved.

Because every unique pair must be evaluated independently, the total number of
assessments scales quadratically as $\binom{N}{2} = N(N-1)/2$.  At an assumed
throughput of 100 pair assessments per person-day, the total workload in
person-days is
\begin{equation}
  D(N) = \left\lceil \frac{N(N-1)}{200} \right\rceil,
  \label{eq:workload}
\end{equation}
where $N$ is the number of non-structure entries under consideration.  The
ceiling function avoids under-estimating the workload: a fraction of a
person-day still requires at least one analyst-day.

\subsection{Composite Priority Scoring}
\label{sec:prioritisation}

To rank regulatory substances by urgency, a composite priority score was
constructed by combining two independent evidence dimensions: regulatory list
membership and intrinsic biological hazard.  The approach is conceptually
grounded in the ToxPi (Toxicological Priority Index) framework~\cite{Reif2010},
which integrates heterogeneous evidence streams into a single prioritisation
score through transparent, domain-specific weights.  The input is the flat CSV
table produced by the ChEBI pipeline described in Section~\ref{sec:kg}
(\texttt{data/processed/chebi.csv}), which contains, per substance, the set of
ECHA regulatory lists on which it appears and the set of ChEBI biological role
labels assigned to it.

\subsubsection{Domain 1: Regulatory List Weights}
\label{sec:list-weights}

Each regulatory list was assigned a weight on an ordinal scale of~0 to~5,
derived directly from the legal consequences triggered by each instrument under
REACH~\cite{REACH2006}, CLP~\cite{CLP2008}, and the POPs
Regulation~\cite{POPs2019}.  Table~\ref{tab:list-weights} summarises the
assignments.


The list score for a given substance is the sum of the weights of all lists on
which it appears:
\begin{equation}
  \text{list\_score} = \sum_{i} w_{\text{list},i},
  \label{eq:list-score}
\end{equation}
where $w_{\text{list},i}$ is the weight of list~$i$ and the sum runs over all
lists containing the substance.  The numerical scale of~0--5 is not prescribed
by any regulatory framework; it was chosen to reflect the ordinal REACH
hierarchy while remaining interpretable, with the ratio 5:0 (SVHC candidate
list : REACH registration) mirroring the qualitative difference in legal
consequence.

\begin{table}[!ht]
	\caption{\textbf{} Regulatory list weights used in Domain~1 of the
		composite priority score.  The weight hierarchy reflects the legal consequence
		of each instrument under REACH (Arts.\ 57--73), CLP (Annex~VI), and the EU
		POPs Regulation.}
	\label{tab:list-weights}
	\centering
	\footnotesize
	\begin{tabularx}{\linewidth}{@{}c p{3.8cm} X@{}}
		\hline
		\textbf{Weight} & \textbf{List or instrument} & \textbf{Legal basis} \\
		\hline
		5 & Candidate list; SVHC identification & Formal SVHC identification, REACH Arts.\ 57--59; triggers communication obligations (Art.\ 33) \\
		4 & Restriction list; authorisation list; POPs list & Binding restrictions or use prohibition, REACH Arts.\ 60--73; elimination obligation, POPs Reg.\ (EU) 2019/1021 \\
		3 & Restriction/authorisation/POPs process; PBT assessment; ED assessment & Active regulatory scrutiny; PBT/ED assessments correspond directly to SVHC criteria, REACH Arts.\ 57(d)--(f) and Annex~XIII \\
		2 & Harmonised CLP list; CLH process; pesticide approval & Legally binding hazard classifications, CLP Annex~VI; positive safety determination required, Reg.\ (EC) No 1107/2009 \\
		1 & Substance evaluation; dossier evaluation; EU positive list & Regulatory scrutiny without a final hazard determination \\
		0 & REACH registration; anhydrous forms & Baseline administrative requirement; no inherent hazard determination, REACH Title~II \\
		\hline
	\end{tabularx}
\end{table}

\subsubsection{Domain 2: Biological Hazard Weights}
\label{sec:bio-weights}

Biological hazard weights were assigned to ChEBI role labels on a scale of~1
to~3, aligned with the SVHC hazard criteria of REACH Art.~57 and the EU
Chemicals Strategy for Sustainability~\cite{EUCSS2020}.  Only substances with a
known InChIKey can receive a ChEBI biological role annotation
(approximately~21\% of the dataset); substances without a ChEBI match are
scored on regulatory list membership alone.

CMR properties (carcinogenic, genotoxic, mutagenic, and teratogenic agents)
receive a weight of~3, consistent with their equal legal standing as Tier-1
SVHC criteria under REACH Arts.~57(a)--(c).  The ChEBI ontology uses the label
\textit{teratogenic agent} (via \texttt{RO:0000087}) for reproductive and
developmental toxicity, rather than the CLP term ``reproductive toxicant'';
this distinction must be respected when mapping ChEBI roles to REACH hazard
classes.  Persistent organic pollutants and endocrine disruptors also receive a
weight of~3, corresponding to REACH Arts.~57(d)--(f) and the elevated priority
assigned to endocrine disruption by the EU CSS~\cite{EUCSS2020}.

Neurotoxins, ecotoxins, and confirmed allergens (sensitisation category~1A)
receive a weight of~2.  Neurotoxicity is identified as a priority hazard class
in the EU CSS~\cite{EUCSS2020}; ecotoxicity is a PBT-adjacent concern under
REACH Annex~XIII.  Neither property is an explicit SVHC criterion under Art.~57,
justifying placement below CMR/POP/ED at weight~2.

Sensitisers (category~1B), antimicrobial agents, and pesticidal activities
(herbicide, insecticide, fungicide, acaricide, nematocide) receive a weight of~1,
reflecting biological interaction with living systems without reaching the
threshold of systemic concern associated with higher tiers.

The biological hazard score is the sum of the weights of all ChEBI roles
associated with a substance:
\begin{equation}
  \text{bio\_score} = \sum_{j} w_{\text{bio},j},
  \label{eq:bio-score}
\end{equation}
where $w_{\text{bio},j}$ is the weight of biological role~$j$.

\subsubsection{Composite Score and Aggregation}
\label{sec:composite-score}

The composite priority score combines both domains by additive summation:
\begin{equation}
  \text{priority\_score} = \text{list\_score} + \text{bio\_score}.
  \label{eq:priority-score}
\end{equation}
Whereas ToxPi computes a weighted average of normalised domain scores (each
rescaled to $[0,1]$)~\cite{Reif2010}, additive summation was preferred here
because the two domains are complementary and largely independent: a substance
may appear on many regulatory lists but lack a ChEBI InChIKey match, or
conversely carry strong biological hazard evidence whilst appearing on few lists.
Normalising one domain against the other would suppress the contribution of the
domain with the smaller raw value, reducing interpretability.  As a consequence
the score is unbounded and is interpreted in relative rather than absolute terms.

\subsubsection{Visualisations}
\label{sec:priority-vis}

Seven complementary figures were produced to answer distinct questions about the
priority landscape:

\begin{enumerate}
  \item A ranked bar chart of the top-50 substances by composite score, identifying
    which substances are most urgent overall.
  \item A heatmap of ECHA list membership for the top-50 substances, revealing
    which regulatory instruments drive each substance's score.
  \item An UpSet plot restricted to concern-level lists, showing intersection
    patterns among the highest-weight instruments.
  \item A bar chart of hazard-type prevalence, broken down by evidence domain,
    illustrating which hazard categories most often contribute to the bio-hazard
    score.
  \item A bubble chart of regulatory breadth (number of lists) against bio-hazard
    score, partitioned into four quadrants using the median number of list
    memberships as the horizontal threshold and $\text{bio\_score} > 0$ as the
    vertical threshold.  The bio-score threshold was fixed at zero rather than
    the median because the bio-score distribution is zero-inflated (approximately
    79\% of substances lack a ChEBI match), which would produce a degenerate
    two-quadrant split at the median.  The quadrants identify: (Q1) broad
    regulatory coverage and biological hazard evidence --- the most urgent
    substances; (Q2) limited regulatory coverage but biological hazard
    evidence --- potential regulatory gaps; (Q3) broad regulatory coverage but
    no biological hazard evidence --- substances requiring structural
    identification to enable evidence-based assessment; and (Q4) limited
    coverage and no biological hazard evidence.
  \item An alluvial diagram tracing the flow from hazard type to regulatory
    instrument, revealing whether the regulatory system has captured each hazard
    class.
  \item A treemap in which tile area represents the number of substances per
    chemical class and tile colour represents the mean priority score, enabling
    identification of common but under-prioritised chemical classes.
\end{enumerate}
