\section{Results}
\label{sec:results}

\subsection{Dataset Overview}

Integration of 18 regulatory sources yielded 41,813 records. Of these,
63\% could be linked to a defined molecular structure via InChIKey.

\subsection{Heterogeneity of Substance Definitions}

A substantial fraction of regulatory entries does not correspond to a
well-defined chemical structure. Approximately 37\% of entries are
non-structure substances, including mixtures (UVCBs), substance groups,
analytical parameters, or entries lacking usable identifiers.

This heterogeneity varies strongly between regulatory instruments:
some lists are largely structure-based, while others are dominated by
group or mixture definitions.

\subsection{Limitations of CAS Numbers}

Although CAS numbers are widely used, they do not provide a reliable
unique identifier. Multiple CAS numbers may refer to the same structure,
and conversely, a single CAS number may correspond to multiple structures.
This confirms that CAS-based linking is insufficient for interoperability.

\subsection{Limited Interoperability Across Regulatory Lists}

Cross-list analysis shows that most substances are unique to a single
regulatory source. Overlap between lists is limited, reflecting the
domain-specific design of regulatory instruments rather than a shared
chemical reference framework.

\subsection{Failure of Semantic Approaches for Non-Structure Substances}

Embedding-based clustering of non-structure substance names yields
semantically coherent groups, but does not resolve chemical identity.
Matching to chemical taxonomies (ChemOnt) succeeds for less than 1\%
of cases with high confidence.

These results indicate that semantic methods cannot compensate for the
absence of structure-based identifiers.

\subsection{Knowledge Graph Integration}

Structure-defined substances can be successfully integrated across
regulatory lists, chemical taxonomies (ChemOnt), and biological
annotations (ChEBI) using InChIKey as a linking key. This demonstrates
the feasibility of a linked-data approach when structure information is
available.

\subsection{Scale of the Interoperability Problem}

For the 15,000+ non-structure entries, full manual reconciliation would
require on the order of one million pairwise comparisons, corresponding
to several thousand person-years of expert effort.

This makes retrospective harmonisation infeasible at scale.

\subsection{Priority Patterns}

Combining regulatory presence with biological hazard annotations enables
identification of high-priority substances and chemical classes.
A small number of chemical classes accounts for a large fraction of
regulated substances, suggesting that class-level approaches can support
regulatory alignment.