\section{Conclusion}
\label{sec:conclusion}
In this study, we show that interoperability of chemical substance data in European regulatory frameworks is fundamentally limited by the lack of a shared and formalised notion of chemical identity. Analysis of 18~regulatory datasets reveals that a substantial fraction of substances cannot be unambiguously linked to a molecular structure, and that commonly used identifiers such as CAS numbers are insufficient for reliable and consistent cross-dataset integration.

The three research questions are answered as follows. \hyperref[sec:question1]{RQ1}: A seven-tier linkability taxonomy (Section~\ref{sec:linkability}) applied to 41,813 records shows that 62.5\% of regulatory entries (20,909) resolve directly via InChIKey (tier~1), while the remaining 37.5\% cannot be linked to a defined chemical structure: 11.6\% carry unresolvable CAS numbers (tier~4), 10.8\% are mixtures or UVCBs (tier~5), and 14.9\% lack any usable identifier (tier~7). \hyperref[sec:question2]{RQ2}: CAS numbers prove unreliable as unique identifiers; bidirectional mapping shows that 195 InChIKeys are each linked to more than one CAS number (some carrying up to seven) and 4 CAS numbers resolve to two distinct InChIKeys, with inconsistencies arising from differences in stereochemistry, salt forms, and registration context. Cross-list analysis further shows that 70\% of structure-resolved InChIKeys appear in only a single regulatory source, reflecting domain-specific rather than shared chemical reference frameworks. \hyperref[sec:question3]{RQ3}: Semantic and embedding-based approaches cannot compensate for the absence of structure-based identifiers; the best-performing model achieves a ChemOnt Hit@1 of only 5.75\%, meaning the correct chemical class is retrieved in fewer than 1 in 17 attempts, and all other models scored below Hit@1~$= 1\%$. Manual inspection reveals that high cosine similarity scores reflect surface-form lexical overlap rather than chemical understanding, a failure inherent to the data: textual descriptions of mixtures, substance groups, and analytical parameters simply do not contain sufficient information to infer molecular structure.

These limitations directly affect semantic interoperability. Many regulatory definitions lack explicit structure-based identifiers and formal semantics, leaving relationships such as equivalence, inclusion, or overlap implicit. This is particularly evident for substance groups and analytical parameters, which often represent sets or transformations of underlying chemical entities (Section~\ref{sec:semantic_mismatch}). Without explicit formalisation of these relationships, interoperability and automated reasoning remain severely constrained.

The scale of the problem becomes clear when considering the effort required for manual reconciliation. For the more than 11,000 non-structure entries, establishing semantic relationships requires pairwise comparison, leading to quadratic growth in curation workload. Under conservative assumptions this corresponds to approximately 630,000 person-days (~3,150 person-years) of expert effort, demonstrating that retrospective harmonisation is not operationally feasible. In contrast, structure-defined substances can be successfully integrated at scale: the knowledge graph constructed in this study (Section~\ref{sec:knowledge_graph}) links regulatory entities to InChIKey, ChemOnt, and ChEBI annotations across approximately 46,600 cross-domain links, enabling cross-domain queries and composite priority scoring (Section~\ref{sec:priority_score}) that are infeasible on fragmented regulatory datasets.

Our results indicate that interoperability cannot be achieved through post hoc technical solutions alone. Instead, it requires changes at the level of data modelling and regulation: the inclusion of structure-based identifiers~(e.g.~InChIKeys), explicit formal semantic models for substance groups and mixtures, and publication of regulatory data as linked data with persistent, resolvable URIs. Without such changes, fragmentation between regulatory and scientific data infrastructures will persist.

Future work should address several open directions. First, formal semantic representations are needed for mixtures, UVCBs, and substance groups that can express their compositional and set-theoretic structure in a machine-readable way, for example representing a substance group as an explicit set of member InChIKeys, or an analytical sum parameter as a function over named constituents. Second, these representations must be translated into operational standards for regulatory data exchange, embedding interoperability requirements directly into the design of future regulatory instruments and reporting obligations. Third, the generality of the findings should be assessed beyond European ECHA datasets, across other regulatory domains, jurisdictions~(e.g.~US EPA and UK HSE), and substance categories where similar semantic mismatches are likely present. Fourth, the 37.5\% of non-structure entries that currently resist automated linking remain an open challenge: curated resolution workflows, possibly guided by the linkability taxonomy introduced here, could provide a principled starting point for targeted harmonisation. Finally, the knowledge graph infrastructure developed as part of this study provides a foundation for further work on automated reasoning, regulatory impact assessment, and cross-domain prioritisation, capabilities that depend critically on the continued expansion of structure-grounded and ontology-linked regulatory data.